---
title: "Home Affairs Variable Code Evidence Register"
subtitle: "Generated DHA/Home Affairs variables mapped to public sources, PLIDA metadata, and not-found evidence gaps"
date: "2026-06-16"
format:
  html:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

## Purpose

This register is the variable-level source audit for generated Department of
Home Affairs products: `AMEP`, `MT_DEMOGS`, `SDB`, `TRAVELLERS`, and `VISA`.

The companion CSV,
[`home-affairs-variable-code-evidence.csv`](home-affairs-variable-code-evidence.csv),
has one row per generated variable. It records the observed generated type and
domain/distribution, the local PLIDA metadata or crosswalk match, the public
source used where found, and whether the current generated values are
source-equivalent, concept-only, or local synthetic evidence.

## Use Rule

Use `observed_generated_domain_or_distribution` as the current fplida schema
surface only. Do not use it as an official distribution unless
`evidence_status` and `source_summary` say that an official source supports the
same value domain.

Rows marked `not_found`, `unverified`, `not_public`, `not_source_equivalent`,
or `distribution_local` are explicit blockers for future source-equivalent
FPLIDA builds.

```{r}
evidence <- read.csv("home-affairs-variable-code-evidence.csv",
                     stringsAsFactors = FALSE,
                     check.names = FALSE)
```

## Coverage

```{r}
as.data.frame(table(evidence$dataset)) |>
  stats::setNames(c("Dataset", "Generated variable rows"))
```

## Evidence Status Counts

```{r}
status_counts <- sort(table(evidence$evidence_status), decreasing = TRUE)
data.frame(
  evidence_status = names(status_counts),
  rows = as.integer(status_counts),
  row.names = NULL
)
```

## Rows Requiring Follow-Up

```{r}
problem_rows <- evidence[grepl("not_found|unverified|not_public|not_source_equivalent|distribution_local",
                               evidence$evidence_status),
                         c("dataset", "table", "variable",
                           "observed_generated_domain_or_distribution",
                           "evidence_status", "source_summary",
                           "required_next_action")]
problem_rows
```

## Source Notes

- SDB rows use Home Affairs settlement reporting pages, reporting caveats, and
  the Settlement Database Data Dictionary. The dictionary supports settlement
  reporting concepts, but many generated compact values still require exact
  table-to-variable mapping before they are source-equivalent.
- AMEP rows use the AMEP profile paper for program, tuition-hours, enrolment,
  and English-proficiency context. No public AMEP variable-level ASLPR
  codebook or generated-score distribution was found in this pass.
- Visa, traveller, and migrant-demographics rows remain mostly local
  PLIDA/generator evidence. No public Home Affairs codebook was found for the
  generated visa subclass/reporting-group labels or the generated traveller
  `1`/`2`/`3` compact quarterly status encoding.

