---
title: "NDIS Variable Code Evidence Register"
subtitle: "Generated NDIS variables mapped to public sources, PLIDA metadata, and not-found evidence gaps"
date: "2026-06-16"
format:
  html:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

## Purpose

This register is the variable-level source audit for generated NDIS participant
variables.

The companion CSV,
[`ndis-variable-code-evidence.csv`](ndis-variable-code-evidence.csv), has one
row per generated variable. It records the observed generated type and
domain/distribution, the local PLIDA metadata or crosswalk match, the public
source used where found, and whether the current generated values are
source-equivalent, concept-only, or local synthetic evidence.

## Use Rule

Use `observed_generated_domain_or_distribution` as the current fplida schema
surface only. Do not use it as an official distribution unless
`evidence_status` and `source_summary` say that an official source supports the
same value domain.

Rows marked `unverified`, `need_mapping`, `not_public`, or
`distribution_local` are explicit blockers for future source-equivalent FPLIDA
builds.

```{r}
evidence <- read.csv("ndis-variable-code-evidence.csv",
                     stringsAsFactors = FALSE,
                     check.names = FALSE)
```

## Coverage

```{r}
as.data.frame(table(evidence$dataset)) |>
  stats::setNames(c("Dataset", "Generated variable rows"))
```

## Evidence Status Counts

```{r}
status_counts <- sort(table(evidence$evidence_status), decreasing = TRUE)
data.frame(
  evidence_status = names(status_counts),
  rows = as.integer(status_counts),
  row.names = NULL
)
```

## Rows Requiring Follow-Up

```{r}
problem_rows <- evidence[grepl("unverified|need_mapping|not_public|distribution_local",
                               evidence$evidence_status),
                         c("dataset", "table", "variable",
                           "observed_generated_domain_or_distribution",
                           "evidence_status", "source_summary",
                           "required_next_action")]
problem_rows
```

## Source Notes

- NDIS dataset pages support participant, plan-budget, utilisation, diagnosis,
  plan-management, support-category, and demographic concepts. They do not, by
  themselves, prove the generated compact categories or plan-budget
  distribution.
- NDIS management-option guidance confirms self-management, plan management,
  and NDIA-management concepts. The generated labels `Self`, `Plan`, and
  `Agency` still need exact DIL/public data-rule mapping before they are
  treated as official labels.

