---
title: "BLADE Dataset Complexity Demo"
subtitle: "Small-sample examples using DIL-backed BLADE tables and local linkage lore"
date: "2026-06-05"
format:
  html:
    toc: true
    number-sections: true
execute:
  eval: true
  message: false
  warning: false
---

## Purpose

This note demonstrates the BLADE behaviours built into the generated dataset
itself. It uses public-DIL-backed BLADE products plus
`_system/plida-blade-link.parquet`, the local fplida person-business link for
PLIDA-to-BLADE exercises:

- mixed lower- and upper-case join variables;
- repeated `spine_id` values after linking person records to business records;
- PAYG business headcounts that do not always reconcile to linked person-job
  rows;
- health-business identification using Table 1 ANZSIC/SISCA fields and Table
  17 occupation fields.

## Build a Small Sample

```{r}
library(arrow)
library(dplyr)

repo_root <- normalizePath(file.path(getwd(), "../.."), mustWork = FALSE)
if (!file.exists(file.path(repo_root, "R", "generate_blade.R"))) {
  repo_root <- normalizePath(getwd(), mustWork = TRUE)
}
if (!file.exists(file.path(repo_root, "R", "generate_blade.R"))) {
  stop("Run this document from the fplida source checkout.")
}
options(fplida.blade_metadata_dir = file.path(repo_root, "inst", "blade_metadata"))

source(file.path(repo_root, "R", "linkage.R"))
source(file.path(repo_root, "R", "export.R"))
source(file.path(repo_root, "R", "generate_dil_lightweight.R"))
source(file.path(repo_root, "R", "generate_blade.R"))

tmp <- file.path(tempdir(), "fplida-blade-complexities-demo")
unlink(tmp, recursive = TRUE)
dir.create(tmp, recursive = TRUE)

run_dir <- file.path(tmp, "fplida_demo")
dir.create(file.path(run_dir, "_system"), recursive = TRUE)
options(fplida.run_dir = run_dir)

n <- 180L
spine <- data.frame(
  spine_id = sprintf("SP%010d", seq_len(n)),
  aeuid_ato = sprintf("ATO%09d", seq_len(n)),
  aeuid_abs = sprintf("ABS%09d", seq_len(n)),
  aeuid_dhda = sprintf("DHDA%08d", seq_len(n)),
  birth_year = rep(1954:2003, length.out = n),
  sex = rep(1:2, length.out = n),
  state = rep(1:8, length.out = n),
  industry = rep(1:19, length.out = n),
  sector = rep(1:4, length.out = n),
  baseline_employed = as.integer(seq_len(n) %% 4L != 0L),
  baseline_income = round(42000 + (seq_len(n) %% 80L) * 1100, 2),
  anzsco_code = rep(c("221111", "241411", "321211", "351411", "552211"),
                    length.out = n),
  anzsco_title = rep(c("Accountant", "Secondary School Teacher",
                       "Motor Mechanic", "Cook", "Bank Worker"),
                     length.out = n),
  anzsco_major = rep(c(2L, 2L, 3L, 3L, 5L), length.out = n),
  stringsAsFactors = FALSE
)
spine$baseline_income[spine$baseline_employed == 0L] <- 0

business_spine <- generate_blade_business_spine(
  spine = spine,
  seed = 42L,
  n_businesses = 46L,
  output_dir = tmp,
  return_data = TRUE,
  overwrite = TRUE
)

generate_blade(
  business_spine = business_spine,
  spine = spine,
  seed = 42L,
  tables = c(1L, 5L, 7L, 17L),
  output_dir = tmp,
  max_rows = 250L,
  include_keys = TRUE
)

blade_dir <- file.path(run_dir, "abs-blade")
table1 <- read_parquet(file.path(
  blade_dir, "blade-table-01-cross-sectional-indicative-data-items.parquet"
))
payg <- read_parquet(file.path(
  blade_dir, "blade-table-05-pay-as-you-go-payg.parquet"
))
stp <- read_parquet(file.path(
  blade_dir, "blade-table-07-single-touch-payroll-stp.parquet"
))
eeh <- read_parquet(file.path(
  blade_dir, "blade-table-17-employee-earning-and-hours-eeh.parquet"
))
link <- read_parquet(file.path(run_dir, "_system", "plida-blade-link.parquet"))

tibble(
  product = c("Table 1 classification", "Table 5 PAYG",
              "Table 7 STP", "Table 17 EEH", "PLIDA-BLADE link"),
  rows = c(nrow(table1), nrow(payg), nrow(stp), nrow(eeh), nrow(link))
)
```

## Naming Normalisation

Table 1 uses lower-case `bn`; the local PLIDA-BLADE link includes upper-case
`BN` aliases alongside the lower-case BLADE key.

```{r}
link_upper <- link |>
  select(SPINE_ID = spine_id, SYNTHETIC_AEUID_ATO, BN,
         synthetic_aeuid_abs, ANZSCO_CODE)

tryCatch(
  left_join(link_upper, table1, by = "bn"),
  error = function(e) conditionMessage(e)
)
```

The intended correction is to declare the asymmetric key or normalise names.

```{r}
link_with_business <- link_upper |>
  left_join(table1, by = c("BN" = "bn"))

link_with_business |>
  select(SPINE_ID, SYNTHETIC_AEUID_ATO, BN, id, d_div06, x_anzsic06,
         x_sisca08, x_sector, ANZSCO_CODE) |>
  head(8)
```

## Duplicate `spine_id` Rows

The local PLIDA-BLADE link is a person-business relationship file, not a
unique person spine. Secondary jobs and owner relationships mean a person can
appear more than once.

```{r}
link |>
  count(spine_id, name = "rows_after_link") |>
  filter(rows_after_link > 1L) |>
  arrange(desc(rows_after_link), spine_id) |>
  head(10)
```

## PAYG Headcount Reconciliation

PAYG Table 5 reports a BN-level `hcnt`. The linked person-job rows provide a
separate count of distinct linked ATO persons. These do not always agree.

```{r}
linked_payg <- link |>
  filter(relationship_type %in% c("employee", "employee_secondary_job")) |>
  group_by(BN) |>
  summarise(
    linked_rows = n(),
    linked_distinct_persons = n_distinct(SYNTHETIC_AEUID),
    .groups = "drop"
  )

payg_reconciliation <- payg |>
  left_join(linked_payg, by = c("bn" = "BN")) |>
  mutate(
    linked_rows = coalesce(linked_rows, 0L),
    linked_distinct_persons = coalesce(linked_distinct_persons, 0L),
    hcnt_gap = hcnt - linked_distinct_persons
  )

payg_reconciliation |>
  filter(hcnt_gap != 0L) |>
  arrange(desc(abs(hcnt_gap))) |>
  head(10)
```

Table 7 STP has its own `d_total_payees`, so it can be compared directly to
PAYG `hcnt`.

```{r}
stp |>
  select(bn, d_total_payees) |>
  left_join(payg |> select(bn, hcnt), by = "bn") |>
  mutate(payee_gap = hcnt - d_total_payees) |>
  filter(payee_gap != 0L) |>
  head(10)
```

## Health Businesses

Table 1 identifies health businesses through ANZSIC Division Q and SISCA/sector
fields. Table 17 provides employee occupation groups, so health businesses can
be triangulated from both business industry and worker occupations.

```{r}
health_businesses <- table1 |>
  filter(d_div06 == "Q") |>
  select(id, bn, d_div06, x_anzsic06, x_sisca08, x_sector)

health_occupations <- eeh |>
  mutate(health_occ = substr(anzsco22_4_eeh, 1L, 2L) %in%
           c("25", "41", "42")) |>
  group_by(id) |>
  summarise(
    employee_rows = n(),
    health_worker_rows = sum(health_occ, na.rm = TRUE),
    .groups = "drop"
  )

health_businesses |>
  left_join(health_occupations, by = "id") |>
  arrange(desc(health_worker_rows), id) |>
  head(10)
```
