---
title: "DHDA Health Variable Code Evidence Register"
subtitle: "Generated Health variables mapped to public sources, PLIDA metadata, and not-found evidence gaps"
date: "2026-06-16"
format:
  html:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

## Purpose

This register is the variable-level source audit for generated DHDA/Health
products: `MBS`, `PBS`, `AIR`, `MCD`, and `NACDC`.

The companion CSV,
[`dhda-health-variable-code-evidence.csv`](dhda-health-variable-code-evidence.csv),
has one row per generated variable. It records:

- the actual generated type, small-domain values, or sampled continuous/date
  distribution;
- the matched PLIDA metadata or concept crosswalk;
- the public source used, if found;
- whether the generated values are source-equivalent, concept-only, or local
  synthetic evidence.

## Use Rule

Use `observed_generated_domain_or_distribution` as the current fplida schema
surface only. Do not use it as an official distribution unless
`evidence_status` and `source_summary` say that an official source supports the
same value domain.

Rows marked `not_found`, `not_source_equivalent`, `needs_tightening`, or
`distribution_local` are explicit blockers for future source-equivalent FPLIDA
builds.

```{r}
evidence <- read.csv("dhda-health-variable-code-evidence.csv",
                     stringsAsFactors = FALSE,
                     check.names = FALSE)
```

## Coverage

```{r}
as.data.frame(table(evidence$dataset)) |>
  stats::setNames(c("Dataset", "Generated variable rows"))
```

## Evidence Status Counts

```{r}
status_counts <- sort(table(evidence$evidence_status), decreasing = TRUE)
data.frame(
  evidence_status = names(status_counts),
  rows = as.integer(status_counts),
  row.names = NULL
)
```

## Not Found or Not Source-Equivalent Rows

```{r}
problem_rows <- evidence[grepl("not_found|not_source_equivalent|needs_tightening",
                               evidence$evidence_status),
                         c("dataset", "variable",
                           "observed_generated_domain_or_distribution",
                           "evidence_status", "source_summary",
                           "required_next_action")]
problem_rows
```

## Source Notes

- MBS rows use ABS PLIDA data-inclusion context, AIHW MBS collection context,
  MBS Online downloads, MBS classification specifications, and local PLIDA
  variable metadata. Generated amount/frequency distributions are local unless
  separately calibrated.
- PBS rows use ABS PLIDA data-inclusion context, the PBS API V3 data
  dictionary, PBS fees/co-payment notes, PBS expenditure reports, and local
  PLIDA variable metadata. Generated amount/frequency distributions are local
  unless separately calibrated.
- AIR rows use ABS PLIDA context and Department of Health AIR reporting
  guidance. No public AIR extract code table was found for the compact antigen,
  episode-status, or schedule-classification values.
- MCD rows use ABS PLIDA context, Services Australia statistical-data context,
  and public Consumer Directory context. No public generated-variable-level MCD
  codebook was found in this pass.
- NACDC rows use ABS PLIDA context plus AIHW/GEN NACDC sources and the May 2026
  table specifications workbook. Several generated compact values are
  deliberately marked concept-only or not source-equivalent.

