---
title: "Education Variable Code Evidence Register"
subtitle: "Generated HE and AEDC variables mapped to public sources, PLIDA metadata, and not-found evidence gaps"
date: "2026-06-16"
format:
  html:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

## Purpose

This register is the variable-level source audit for generated Department of
Education products: `HE` and `AEDC`.

The companion CSV,
[`education-variable-code-evidence.csv`](education-variable-code-evidence.csv),
has one row per generated variable. It records the observed generated type and
domain/distribution, the local PLIDA metadata match, the public source used
where found, and whether the current generated values are source-equivalent,
concept-only, or local synthetic evidence.

## Use Rule

Use `observed_generated_domain_or_distribution` as the current fplida schema
surface only. Do not use it as an official distribution unless
`evidence_status` and `source_summary` say that an official source supports the
same value domain.

Rows marked `not_exhausted`, `cutoffs_local`, `distribution_local`, or
`not_public` are explicit blockers for future source-equivalent FPLIDA builds.

```{r}
evidence <- read.csv("education-variable-code-evidence.csv",
                     stringsAsFactors = FALSE,
                     check.names = FALSE)
```

## Coverage

```{r}
as.data.frame(table(evidence$dataset)) |>
  stats::setNames(c("Dataset", "Generated variable rows"))
```

## Evidence Status Counts

```{r}
status_counts <- sort(table(evidence$evidence_status), decreasing = TRUE)
data.frame(
  evidence_status = names(status_counts),
  rows = as.integer(status_counts),
  row.names = NULL
)
```

## Rows Requiring Follow-Up

```{r}
problem_rows <- evidence[grepl("not_exhausted|cutoffs_local|distribution_local|not_public",
                               evidence$evidence_status),
                         c("dataset", "table", "variable",
                           "observed_generated_domain_or_distribution",
                           "evidence_status", "source_summary",
                           "required_next_action")]
problem_rows
```

## Source Notes

- HE rows use the Department of Education student-data page and TCSI data
  element pages. Attendance type and mode have exact public TCSI anchors in
  this pass; most other generated HE category codes still need row-level TCSI
  element mapping before labels are treated as official.
- AEDC rows use the Department of Education AEDC page and AEDC 2025 Data
  Dictionary. The dictionary is a public variable reference, but generated
  scores, vulnerability categories, and `DV1`/`DV2` cutoffs are still local
  until checked against official scoring/reference tables.

